Enhance Road Detection Data Processing of LiDAR Point Clouds to Specifically Identify Unmarked Gravel Rural Roads

### 2023-05-02 Huston_Paper.tex
formatting + moar stuff

	\begin{figure}[H]
		\centering
		\begin{subfigure}{0.45\textwidth}
			\centering
			\includegraphics[width=1.0\linewidth,height=5.0 cm,keepaspectratio]{figures/vlcsnap-2023-04-20-08h48m17s447}
			\caption[Blackburn Road Camera View]{}
			\label{fig:Bean_Cam_View}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
			\centering
			\includegraphics[width=1.0\linewidth,height=5.0 cm,keepaspectratio]{figures/gravel_lot_pic.png}
			\caption[Armig Road Camera View]{}
			\label{fig:Armig_Road_Camera_View}
		\end{subfigure}
		\caption[Blackburn Road \& Gravel Parking Lot]{Road view of Blackburn Road - an unmarked asphalt road (a). Road view of the gravel driveway (b).}
		\label{fig:Combined_Roads}
	\end{figure}

@@

%   (Figure \ref{fig:flowzchart})
%	\begin{figure}
%		\centering
%		\includegraphics[width=0.9\linewidth]{figures/flowz}
%		\caption[Project Flow]{Project flowchart}
%		\label{fig:flowzchart}
%	\end{figure}

	% Quadrant size inversely affects rate of classification, as smaller quadrants allow for greater resolution at the expense of more computational time.

%	{Testing consecutive scans required a method of compiling LiDAR data into a single point cloud and manually classifying areas for road detection accuracy. This was completed by post processing rosbags that contained scanning LiDAR, GPS, and IMU data of an unmarked road. Point cloud translation and rotation was accomplished using transformation matrices derived from GPS and IMU data. Rotation matrices were created using extracted roll, pitch, and yaw data from the IMU. LiDAR origin was found using GPS longitude, latitude, and altitude data. Physical distances between the GPS, IMU, and LiDAR was rectitude by obtaining reference frames provided by AutonomouStuff. GPS coordinates were offset by the current orientation and converts the ground truth to the LiDAR frame. GPS, IMU, and LiDAR reference frames and rotational updates were combined into trajectory vectors. Consecutive point cloud scans may then translated and rotated with derived transformation matrices (Figure \ref{fig:Compiled_PCD}). While not as robust as more sophisticated methods of point cloud aggregation such as NDT or ICP scan matching, over shorter distances using this method proved to be adequate for scoring purposes. Six stretches of unmarked roads were examined for accuracy: three of chipseal and three of gravel. Road and side-of-road areas were manually defined and projected unto consecutive LiDAR scans to score classification results. Due to the fuzzy nature of the classification results, the side-of-road areas were roughly a standard lane width to the sides of the road to accommodate the inability to define a rigid road edge.}

@@

%	{ Comparing averaged X, Y, and Z coordinates of the classified quadrant allowed for easier comparison and scoring of the RDF classification to the manually classified areas. 
% Insert sentence or two on rate of of classification, feat extract, etc.; reference figure with rates> This was tested with a machine with 64 GB of DDR4 RAM and a Ryzen 9 5900X 12 core x 24 thread CPU.}

@@

	{Due to the side of the road was being re-grassed near area $3$, the algorithm sometimes struggled to differentiate the dirt from gravel, leading to some miss-classification of the side-of-road areas (Figure \ref{fig:pre-adjust} vs Figure \ref{fig:post-adjust}). Exploitation of confidence scores to adjust results by requiring a very high confidence for positive gravel surface classification was explored. All gravel surfaces that did not meet a 80\% confidence score was re-labeled as "unknown", however these adjustment had either a negligible or detrimental impact on the final results. Overall it was found that an average true-positive accuracy of all intercepting gravel surfaces was 67.54\%. If any asphalt miss-classification on the gravel areas is included, the average true-positive road detection accuracy increases to 76.97\%. It was found that an average true-positive accuracy of all asphalt surfaces was 87.05\%. Differentiation between gravel road surfaces and surrounding terrain is necessary to determine boundaries for trajectory tracking and navigation. Side-of-road areas were established (Figure \ref{fig:prepostadjust}), and the overall average for "unknown" terrain type was found to be 77.49\%.}

@@

	\begin{figure}[H]
		\centering
		\begin{subfigure}{0.495\textwidth}
			\centering
			\includegraphics[width=0.975\linewidth]{figures/side_of_road_nums_4_with_no_filts_2}
			\caption[Pre-Adjusted]{}
			\label{fig:pre-adjust}	
		\end{subfigure}
		\begin{subfigure}{0.495\textwidth}
			\centering
			\includegraphics[width=0.975\linewidth]{figures/side_of_road_nums_4_with_filts_2}
			\caption[Post-Adjusted]{}
			\label{fig:post-adjust}
		\end{subfigure}
		\label{fig:prepostadjust}
		\caption[Post-Adjusted]{Raw classification result example (a). Results were slightly altered by considering confidence scores and inter-arc distances to decrease miss-classification of gravel surfaces, however the effect was negligible (b).}
	\end{figure}

@@

>>>>

	\begin{figure}[H]
		\centering
		\begin{subfigure}{0.45\textwidth}
			\centering
			\includegraphics[width=1.0\linewidth,height=5.0 cm,keepaspectratio]{figures/vlcsnap-2023-04-20-08h48m17s447}
			\caption[Blackburn Road Camera View]{}
			\label{fig:Blackburn_Road_View}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
			\centering
			\includegraphics[width=1.0\linewidth,height=5.0 cm,keepaspectratio]{figures/gravel_lot_pic.png}
			\caption[Armig Road Camera View]{}
			\label{fig:Gravel_Lot_View}
		\end{subfigure}
		\caption[Blackburn Road \& Gravel Parking Lot]{Road view of Blackburn Road - an unmarked asphalt road (a). Road view of the gravel driveway (b).}
		\label{fig:Combined_Roads}
	\end{figure}

@@

@@

@@

	{Overall it was found that an average true-positive accuracy of all intercepting gravel surfaces was 67.54\%. If any asphalt miss-classification on the gravel areas is included, the average true-positive road detection accuracy increases to 76.97\%. It was found that an average true-positive accuracy of all asphalt surfaces was 87.05\%. Differentiation between gravel road surfaces and surrounding terrain is necessary to determine boundaries for trajectory tracking and navigation. Side-of-road areas were established (Figure \ref{fig:prepostadjust}), and the overall average for "unknown" terrain type was found to be 77.49\%.}

@@

%	\begin{figure}[H]
%		\centering
%		\begin{subfigure}{0.495\textwidth}
%			\centering
%			\includegraphics[width=0.975\linewidth]{figures/side_of_road_nums_4_with_no_filts_2}
%			\caption[Pre-Adjusted]{}
%			\label{fig:pre-adjust}	
%		\end{subfigure}
%		\begin{subfigure}{0.495\textwidth}
%			\centering
%			\includegraphics[width=0.975\linewidth]{figures/side_of_road_nums_4_with_filts_2}
%			\caption[Post-Adjusted]{}
%			\label{fig:post-adjust}
%		\end{subfigure}
%		\label{fig:prepostadjust}
%		\caption[Post-Adjusted]{Raw classification result example (a). Results were slightly altered by considering confidence scores and inter-arc distances to decrease miss-classification of gravel surfaces, however the effect was negligible (b).}
%	\end{figure}

	\begin{figure}[H]
			\centering
			\includegraphics[width=0.975\linewidth]{figures/side_of_road_nums_4_with_filts_2}
			\caption[Post-Adjusted]{}
			\label{fig:prepostadjust}
		\caption[Area Scores]{Per-area scores were calculated, indicating algorithm efficiency of unmarked road detection. }
	\end{figure}

@@

	{Classification of real-world data outside of the training data set presents difficulties to the generated Random Decision Forest Classifiers generated in this work. Due to the side of the road was being re-grassed near area $3$ (Figure \ref{fig:area_percentages}, Figure \ref{fig:Blackburn_Road_View}), the algorithm struggled to differentiate the dirt from gravel, leading to some miss-classification of the side-of-road areas. Examination of the feature distribution shows a high amount of overlap (Figure \ref{}). Exploitation of confidence scores to adjust results by requiring a very high confidence for positive gravel surface classification was explored. All gravel surfaces that did not meet a 80\% confidence score was re-labeled as "unknown", however these adjustment had either a negligible or detrimental impact on the final results. Higher confidence score requirements resulted in "unknown" over-classification, while too low resulted in "gravel" over-classification.  }
